---
title: "Coverage Simulation to Compare BYM implementations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r knit_init, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What this vignette explains

This vignette is supplementary material for the article "Coverage of Effect Estimates of Besag-York-Mollié Algorithm in Up-To-Date R Implementations".

We publish our source code both to enable you to replicate our research and to give examples of the function calls for further application.


## Loading required dependencies
```{r setup}
  library("compareBYMimplementations")
  library("rstan")
  library("tidyverse")
```



## Generating a list of all simulation parameter combinations
```{r param}
load(file.path(find.package("compareBYMimplementations"), "data", "shape_GER.Rdata"))

basic_round <- do.call(what = rbind,
                       args = list(data.frame(region = "3",
                                              seed_risk = 1000+1:109),
                                   data.frame(region = "9",
                                              seed_risk = 1000+1:52),
                                   data.frame(region = "D",
                                              seed_risk = 1000+1:14)))
basic_round$region_ID <- 1:nrow(basic_round)


sim_param <- expand.grid(region_ID = basic_round$region_ID,
                         n_convolution = c(2, 4, 6),
                         rate_per_100k = 2^(0:7))

running_df <- dplyr::left_join(x = basic_round,
                               y = sim_param,
                               by = "region_ID")

running_df$seed_poisson <- running_df$seed_risk + 1000
running_df$region <- as.character(running_df$region)
running_df$simulation_ID <- 1:nrow(running_df)

print(as_tibble(running_df[(1:8)*475, ]))
```



## Generating a sample shapefile
... if this was an application on real data, you would submit your own shapefile with data here. For our simulations, we iterated through all rows of the parameter data.frame

```{r shapefile}
    my_running_ID <- 1234

    running_param = list(region        = running_df$region[my_running_ID],
                         ID            = my_running_ID,
                         seed_risk     = running_df$seed_risk[my_running_ID],
                         seed_poisson  = running_df$seed_poisson[my_running_ID],
                         n_convolution = running_df$n_convolution[my_running_ID],
                         rate_per_100k = running_df$rate_per_100k[my_running_ID])

    running_param <- append(x = running_param,
                            values = list(n = switch (running_param$region,
                                                      "9" = 96,
                                                      "3" = 46,
                                                      "D" = 402)))
    states_codes <- switch (running_param$region,
                            "9" = 9,
                            "3" = 3,
                            "D" = 1:16)

    shape_D <- artificial_risk(shapefile     = shape_GER[substr(shape_GER@data$region_ID, 1, 2) %in% sprintf("%02d", states_codes), ],
                               n_convolution = running_param$n_convolution,
                               rate_per_100k = running_param$rate_per_100k,
                               seed_risk     = running_param$seed_risk,
                               seed_poisson  = running_param$seed_poisson)
```


## Prepare Stan for running
```{r stanconf, cache=TRUE}
write(x    = compareBYMimplementations::model_STAN_bym(),
      file = file.path("my_dir/stan",
                       "my_bym.stan"))
write(x    = compareBYMimplementations::model_STAN_bym2(),
      file = file.path("my_dir/stan",
                       "my_bym2.stan"))


options(mc.cores = 2);
rstan_options("auto_write" = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')


bym_stan  = rstan::stan_model(file = file.path("my_dir/stan",
                                               "my_bym.stan"),
                              model_name = "my_bym")
bym2_stan = rstan::stan_model(file = file.path("my_dir/stan",
                                               "my_bym2.stan"),
                              model_name = "my_bym2")
```


## Calling of BYM implementations
we wrote functions smooth_... that unify the function calls of the different implementations.
For infomration on the details of the function calls, we would like to refer to the source code of these functions.

Using these calling functions, the BYM calculations are

```{r callBYM, cache=TRUE, results='hide', echo=TRUE, warning=FALSE}
   res_INLA_bym1 <- smooth_inla(formula      = O ~ E + region_ID,
                                 data         = shape_D@data,
                                 shapefile    = shape_D,
                                 INLA_model   = "bym",
                                 BUGS_tempdir = "my_dir/bugs",
                                 time_scaling = 1)

    res_INLA_bym2 <- smooth_inla(formula      = O ~ E + region_ID,
                                 data         = shape_D@data,
                                 shapefile    = shape_D,
                                 INLA_model   = "bym2",
                                 BUGS_tempdir = "my_dir/bugs",
                                 time_scaling = 1)

    res_BUGS_bym <-  smooth_bugs(formula      = O ~ E + region_ID,
                                 data         = shape_D@data,
                                 shapefile    = shape_D,
                                 INLA_model   = "bym2",
                                 BUGS_tempdir = "my_dir",
                                 time_scaling = 1)
    
    
    res_STAN_bym  <- smooth_stan(formula        = O ~ E + region_ID,
                                 data           = shape_D@data,
                                 shapefile      = shape_D,
                                 STAN_dir       = "my_dir/stan",
                                 STAN_cores     = 2,
                                 STAN_bym_model = bym_stan)
    
    res_STAN_bym2 <- smooth_stan(formula        = O ~ E + region_ID,
                                 data           = shape_D@data,
                                 shapefile      = shape_D,
                                 STAN_dir       = "my_dir/stan",
                                 STAN_cores     = 2,
                                 STAN_bym_model = bym2_stan)
    
    res_CARBayes  <- smooth_rpkg(formula      = O ~ E + region_ID,
                                 data         = shape_D@data,
                                 shapefile    = shape_D,
                                 INLA_model   = "bym2",
                                 STAN_dir     = STAN_dir,
                                 STAN_cores   = 2,
                                 time_scaling = 1)

```

## Summarizing BYM results
We wrote a postestimation function that evaluates the average error and coverage. Calling these, we get
```{r BYMpostest, cache=TRUE}
    a <- summarize_BYMresult(estimation_tbl = shape_D@data,
                             simulation_tbl = res_INLA_bym1$sRR,
                             method_name    = "INLA-BYM",
                             duration       = res_INLA_bym1$duration,
                             region         = running_param$region,
                             n              = running_param$n,
                             ID             = running_param$ID,
                             seed_risk      = running_param$seed_risk,
                             seed_poisson   = running_param$seed_poisson,
                             n_convolution  = running_param$n_convolution,
                             rate_per_100k  = running_param$rate_per_100k)
    b <- summarize_BYMresult(estimation_tbl = shape_D@data,
                             simulation_tbl = res_INLA_bym2$sRR,
                             method_name    = "INLA-BYM2",
                             duration       = res_INLA_bym2$duration,
                             region         = running_param$region,
                             n              = running_param$n,
                             ID             = running_param$ID,
                             seed_risk      = running_param$seed_risk,
                             seed_poisson   = running_param$seed_poisson,
                             n_convolution  = running_param$n_convolution,
                             rate_per_100k  = running_param$rate_per_100k)
    c <- summarize_BYMresult(estimation_tbl = shape_D@data,
                             simulation_tbl = res_BUGS_bym$sRR,
                             method_name    = "OpenBUGS",
                             duration       = res_BUGS_bym$duration,
                             region         = running_param$region,
                             n              = running_param$n,
                             ID             = running_param$ID,
                             seed_risk      = running_param$seed_risk,
                             seed_poisson   = running_param$seed_poisson,
                             n_convolution  = running_param$n_convolution,
                             rate_per_100k  = running_param$rate_per_100k)
  d  <- summarize_BYMresult(estimation_tbl = shape_D@data,
                            simulation_tbl = res_STAN_bym$sRR,
                            method_name    = "Rstan-BYM",
                            duration       = res_STAN_bym$duration,
                            region         = running_param$region,
                            n              = running_param$n,
                            ID             = running_param$ID,
                            seed_risk      = running_param$seed_risk,
                            seed_poisson   = running_param$seed_poisson,
                            n_convolution  = running_param$n_convolution,
                            rate_per_100k  = running_param$rate_per_100k)

  e  <- summarize_BYMresult(estimation_tbl = shape_D@data,
                            simulation_tbl = res_STAN_bym2$sRR,
                            method_name    = "Rstan-BYM2",
                            duration       = res_STAN_bym2$duration,
                            region         = running_param$region,
                            n              = running_param$n,
                            ID             = running_param$ID,
                            seed_risk      = running_param$seed_risk,
                            seed_poisson   = running_param$seed_poisson,
                            n_convolution  = running_param$n_convolution,
                            rate_per_100k  = running_param$rate_per_100k)
    f <- summarize_BYMresult(estimation_tbl = shape_D@data,
                             simulation_tbl = res_CARBayes$sRR,
                             method_name    = "CARBayes",
                             duration       = res_CARBayes$duration,
                             region         = running_param$region,
                             n              = running_param$n,
                             ID             = running_param$ID,
                             seed_risk      = running_param$seed_risk,
                             seed_poisson   = running_param$seed_poisson,
                             n_convolution  = running_param$n_convolution,
                             rate_per_100k  = running_param$rate_per_100k)

    summarized_result <- (do.call(what = rbind,
                                  args = list(a, b, c, d, e, f)))
```

```{r calcout, results = "asis"}
    knitr::kable(t(summarized_result))
```

